---
title: "Software for reckoning AnVIL/Terra usage"
author: "BJ Stubbs, Vincent J. Carey, rebjh/stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Software for reckoning AnVIL/terra usage}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

Detailed billing information is essential for cost and resource planning in cloud based analysis projects, but it can be difficult to obtain. The goal of this software is to help users of the Terra/AnVIL platform get access to this data as easily as possible.

# Setup

The functions in this vignette require a user to connect the billing export in the Google Cloud Platform project associated with Terra/AnVIL to a BigQuery dataset.

General information on this process can be found here:

https://cloud.google.com/billing/docs/how-to/export-data-bigquery

In order to set this up with the AnVIL/Terra system:

1. Create a new project in Google under the billing account linked to the Terra project
2. Create a BigQuery dataset to store the billing export
3. Go to the billing tab in the gcloud console and enable export to this dataset
4. Make sure the user of this software has the BigQuery scope on the billing project

# Obtaining billing data

## Overview

Billing data is generally available 1 day after incurring charges. Billing data is stored in BigQuery in a partitioned table, and is queryable using the bigrquery package.

## Setting up a request

In order to generate a request you need:

1. start: date of start of reckoning
2. end: date of end of reckoning
3. project: GCP project id
4. dataset: GCP dataset id for billing data in BigQuery
5. table: GCP table for billing data in BigQuery

Then you can use the function:

```{}
setup_billing_request(start, end, project, dataset, table, billing_code)
```

To create a request.

Once you have a request object, then you can get the billing data associated with that request using the reckon() function on your billing request.

## output

The result of a reckonging on a billing request is an instance of `avReckoning`

We took a snapshot of usage in a project we work on, and it is available as demo_rec. This request represents one day of usage in AnVIL/Terra.

```{r lkd}
suppressPackageStartupMessages({
library(AnVILBilling)
library(dplyr)
library(magrittr)
})

demo_rec
```
The available keys for the billing object are shown.

For Terra, 3 of the most useful keys are:

1. terra-submission-id : this key is associated with a cromwell workflow execution
2. cromwell-workflow-id : this key is associated with a job in a workflow, so if you apply a workflow to multiple inputs, you will have 1 terra id and multiple cromwell ids
3. goog-dataproc-cluster-name : this key is associated with a jupyter notebook or Rstudio cluster on Terra. The user of these resources can be determined using the AnVIL API R package

```{}
library(AnVIL)
leo=Leonardo()
leo$getCluster(clustername)
```


# Drilling down

Given a key type, we want to know associated values.
```{r lklklk}
v = getValues(demo_rec@reckoning, "terra-submission-id")
v
```

To understand activities associated with this submission,
we subset the table.

```{r adasd}
s = subsetByKeyValue(demo_rec@reckoning, "terra-submission-id", v)
s
```
The following data is available in this object

```{r vars}
names(s)
```

You can drill down more to see what products used during the submission:
```{r lkpr}
AnVILBilling:::getSkus(s)
```
